Thread Pool in GeckoDB/BOLSTER - Implementing a Thread Pool for a High Performance Database System \\

### 2018-07-02 Presentations/Paper/paper_thread_pool.tex
Add first version of conclusion

\begin{abstract}
        Processing tasks in parallel is used in nearly all applications to keep
        up with the requirements of modern software systems. However, the
        current implementation of parallel processing in GeckoDB requires
        spawning many short-lived threads that execute a single task and then
        terminate. This creates a lot of overhead since the threads are not
        being reused. To counter this effect we implemented a thread pool to be
        used in combination with GeckoDB to reduce the additional time required
        to create and close thousands of threads. In this paper, we show our
        implementation of a thread pool to process independent tasks in parallel
        including the waiting for a set of processed tasks. We also compare the
        thread pool implementation against the current one. The evaluation of
        the thread pool shows, that designing and implementing a thread pool is
        a balancing act between generalization of tasks and the processing
        performance. Additionally, we show that the task and thread pool configuration
        depending on the use case has a high impact on the thread
        pool performance. At the end of our evalutation we rate the implementation a
        success because we fulfilled all given constraints and generally reduce
        overhead compared to creating and spawning threads individually.
\end{abstract}

@@

To overcome these drawbacks, we integrate an optimised thread pool in BOLSTER. Along with the implementation, we measure the performance of the primitives to determine the thread pool overhead. Additionally, we measure metrics like \emph{waiting and busy time} of threads to evaluate correct thread pool sizes for the considered use cases. In this work we make the following contributions:
\begin{itemize}
	\item We describe our design and implementation of the thread pool
	\item We evaluate the possibility to wait for a group of task in the calling thread
	\item We compare our thread pool against the existing implementation in BOLSTER
	\item We measure and evaluate statistics for threads, tasks and the thread pool
\end{itemize}
We organized the rest of the paper as follows. In Section 2, we give preliminaries about the considered task configuration and about thread safe access of memory. In Section 3, we show our design and implementation of the thread pool and examine our experimental environment in Section 4. In Section 5, we describe the results of our performance evaluation. In Section 6, we name related work and state our conclusion and future work in Section 7.

@@

\subsection{Processing Tasks}
Processing tasks in a thread pool is a balancing act between generalization the of tasks and the performance of their execution. Therefore, we provide a set of functions to process task in the thread pool to match different requirements of BOLSTER. In Table 1, we show the functions provided by our thread pool to process tasks. One of the requirements of BOLSTER is waiting until a set of tasks is finished. To achieve the same performance as the baseline implementation reaches, we let the main thread first participate in the execution of the tasks and then wait until all tasks with the same handle are finished.

@@

\subsubsection{Evaluation Results}
In Figure \ref{fig7}, we show the relation of waiting and completion time of different thread pool sizes. The waiting time is, compared to the completion time, very small and nearly constant for all considered thread pool sizes. From this observation, we draw the following conclusions for our \emph{first come first serve} scheduling approach. Firstly, constant average waiting time indicates only small performance decreases due to priority queue locking. Secondly, even for $32$ threads in one pool instance, the waiting time occupies only a small fraction of the completion time, which indicates good thread usage.

@@

\section{Conclusion and Future Work}

>>>>

\begin{abstract}
        Processing tasks in parallel is used in nearly all applications to keep
        up with the requirements of modern software systems. However, the
        current implementation of parallel processing in GeckoDB requires
        spawning many short-lived threads that execute a single task and then
        terminate. This creates a lot of overhead since the threads are not
        being reused. To counter this effect we implemented a thread pool to be
        used in combination with GeckoDB to reduce the additional time required
        to create and close thousands of threads. In this paper, we show our
        implementation of a thread pool to process independent tasks in parallel
        including the waiting for a set of processed tasks. We also compare the
        thread pool implementation against the current one. Additionally, we show that the task and thread pool configuration
        depending on the use case has a high impact on the thread
        pool performance. At the end of our evaluation, we show that the implementation 
        fulfils all given constraints and generally reduce
        overhead compared to creating and spawning threads individually.
\end{abstract}

@@

To overcome these drawbacks, we integrate an optimised thread pool in BOLSTER. Along with the implementation, we measure the performance of the primitives to determine the thread pool overhead. Additionally, we measure metrics like \emph{waiting and busy time} of threads to evaluate correct thread pool sizes for the considered use cases. In this work we make the following contributions:
\begin{itemize}
	\item We describe our design and implementation of the thread pool
	\item We evaluate the possibility to wait for a group of task in the calling thread
	\item We compare our thread pool against the existing implementation in BOLSTER
	\item We measure and evaluate additional statistics for different thread pool configurations
\end{itemize}
We organized the rest of the paper as follows. In Section 2, we give preliminaries about the considered task configuration and about thread safe access of memory. In Section 3, we show our design and implementation of the thread pool and examine our experimental environment in Section 4. In Section 5, we describe the results of our performance evaluation. In Section 6, we name related work and state our conclusion and future work in Section 7.

@@

\subsection{Processing Tasks}
Processing tasks in a thread pool is a balancing act between generalization the of tasks and the performance of their execution. With higher generalization, the implementation can be used for more use cases, but the execution performance may decrease. Therefore, we provide a set of functions to process task in the thread pool to match different requirements of BOLSTER. In Table 1, we show the functions provided by our thread pool to process tasks. One of the requirements of BOLSTER is waiting until a set of tasks is finished. To achieve the same performance as the baseline implementation reaches, we let the main thread first participate in the execution of the tasks and then wait until all tasks with the same handle are finished.

@@

\subsubsection{Evaluation Results}
In Figure \ref{fig7}, we show the relation of waiting and completion time of different thread pool sizes. The waiting time is, compared to the completion time, very small and nearly constant for all considered thread pool sizes. From this observation, we draw the following conclusions for our \emph{first come first serve} scheduling approach. Firstly, constant average waiting time indicates only small performance decreases due to priority queue locking. Secondly, even for $32$ threads in one pool instance, the waiting time occupies only a small fraction of the completion time, which indicates good thread usage.  Thirdly, the execution performance increases along with the thread pool size, which shows that even for $32$ threads, parallelization leads to faster task execution.

@@

\section{Conclusion and Future Work}
In this work, we describe our implementation of a thread pool system, present in detail the implementation and show our evaluation results containing a comparison against the baseline implementation and the waiting time behaviour over different thread pool sizes. Within our implementation, we state the differences of parallel memory access as well as the problem of implementing synchronous waiting for a set of tasks.

As presented in the evaluation, our thread pool implementation reaches better execution performance than the baseline because of avoiding the creation of short living threads for each request. We measured the average time threads spent waiting in different thread pool sizes and present, that even for a size of $32$ threads, the waiting time stays constantly small. Consequently, we conclude that using \emph{first come first serve} as scheduling approach leads to small waiting times and high execution performance.

To sum up, our thread pool implementation fulfils all performance and design requirements of geckoDB and BOLSTER and reaches even faster execution time than the existing implementation. In the future, we plan to make the following adaptations to our thread pool system. Firstly, we want to reduce the number of dynamic allocation, since heap allocations mostly are slower than stack allocations. Secondly, we want to examine the usage of existing priority queue implementations for our thread pool system. Thirdly, we plan to integrate a dynamic resizing function based on the collected statistics of the performance monitoring component.
